---
title: "Satimage"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---


```{r echo=F}

library(ggplot2) # For some graphs

library(FactoMineR) # For PCA
library(factoextra) # For PCA
library(caret) # For feature plots
library(e1071) # For SVM
library(nnet) # For Neural nets
library(pROC) # For ROC
library(multiROC) # For multiclass ROC

library(cluster) # For k-means
library(clustertend) # For Hopkins Statistic
library(fpc) # For k-means
library(seriation) # For k-means
library(ggpubr) # For clustering visualization
library(TSclust) # For clustering Evaluation
```

```{r}
# Create a vector containing all attribute names
attribute.names = c()
for(pixel.number in 1:9) {
  for(attribute.suffix in c("Red", "Green", "NIR1", "NIR2")) {
    attribute.name = paste("p", pixel.number, ".", attribute.suffix, sep="")
    attribute.names = c(attribute.names, attribute.name)
  }
}
attribute.names = c(attribute.names, "class")

# Create a vector containing all class names
class.names = c("red soil", "cotton crop", "grey soil", "damp grey soil", "soil with vegetation stubble", "very damp grey soil")

# Load the dataset from the csv file and cast the label to factor
dataset = data.frame(read.csv("dataset.csv", col.names = attribute.names, sep=" "))
dataset$class = factor(dataset$class, labels = class.names)

# Determine the input and the output space
dataset.input = dataset[, 1:ncol(dataset)-1]
dataset.output = dataset$class
```

```{r}
# Determine some statistical measures on the input attributes
dataset.mean = sapply(dataset.input, mean)
dataset.sd = sapply(dataset.input, sd)
dataset.var = sapply(dataset.input, var)
dataset.min = sapply(dataset.input, min)
dataset.max = sapply(dataset.input, max)
dataset.median = sapply(dataset.input, median)
dataset.range = sapply(dataset.input, range)
dataset.quantile = sapply(dataset.input, quantile)

```

```{r}
# Determine class frequencies
class.frequencies = data.frame(table(dataset.output))
colnames(class.frequencies) = c("class", "frequencies")
class.frequencies$percentages = paste(format((class.frequencies$frequencies / sum(class.frequencies$frequencies)) * 100, digits = 2), "%", sep = "")
ggplot(data = class.frequencies, aes(x = class, y = frequencies, fill = class)) + geom_bar(stat = "identity") + geom_text(aes(label = percentages), vjust = 1.5, colour = "white") + theme(axis.text.x=element_blank())
ggplot(data = class.frequencies, aes(x = "", y = frequencies, fill = class)) + geom_bar(stat = "identity", width = 1, color = "white") + coord_polar("y", start = 0)
```
## Separabilità
I valori relativi allo stesso canale tra pixel adiacenti risultano correlati (come evidente dai grafici), questo perche' il dataset e' costruito come una sliding window quadrata 3x3 su una singola immagine satellitare. Ci si aspetta quindi una varianza piuttosto contenuta sui singoli valori e la sopra citata correlazione.

```{r}
featurePlot(dataset.input[, c("p5.Red", "p5.Green", "p5.NIR1")], dataset.output, plot="pairs", scales=list((relation="free"), y=list(relation="free")), auto.key=list(columns=3), main="Feature plot for p5.Red, p5.Green, p5.NIR1 pairs")
featurePlot(dataset.input[, c("p5.Red", "p6.Red", "p4.Red")], dataset.output, plot="pairs", scales=list((relation="free"), y=list(relation="free")), auto.key=list(columns=3), main="Feature plot for p5.Red, p6.Red, p4.Red pairs")
featurePlot(dataset.input[, c("p5.Green", "p6.Green", "p4.Green")], dataset.output, plot="pairs", scales=list((relation="free"), y=list(relation="free")), auto.key=list(columns=3), main="Feature plot for p5.Green, p6.Green, p4.Green pairs")
featurePlot(dataset.input[, c("p5.Green", "p2.Green", "p8.Green")], dataset.output, plot="pairs", scales=list((relation="free"), y=list(relation="free")), auto.key=list(columns=3), main="Feature plot for p5.Green, p2.Green, p8.Green pairs")
featurePlot(dataset.input[, c("p5.Red", "p2.Red", "p8.Red")], dataset.output, plot="pairs", scales=list((relation="free"), y=list(relation="free")), auto.key=list(columns=3), main="Feature plot for p5.Red, p2.Red, p8.Red pairs")
featurePlot(dataset.input[, c("p5.NIR1", "p2.NIR1", "p8.NIR1")], dataset.output, plot="pairs", scales=list((relation="free"), y=list(relation="free")), auto.key=list(columns=3), main="Feature plot for p5.NIR1, p2.NIR1, p8.NIR1 pairs")
```

## PCA
Data la correlazione rilevata precedentemente e l'alto numero di attributi in input sembra più che opportuno eseguire una PCA sul dataset al fine di ridurre la dimensione dello spazio di input.
```{r}
# Execute the PCA. Extract only the first four principal components
pca.results <- PCA(dataset.input, scale.unit = TRUE, ncp = 4, graph = FALSE)

# Get the eigenvalues and plot them. The first four components explain about 92% of variance
pca.results.eig.val <- get_eigenvalue(pca.results)
fviz_eig(pca.results, addlabels = TRUE, ylim = c(0, 50))

# Show the correlation between the variables using the first two principal components
# Many variable are positively correlated
fviz_pca_var(pca.results, col.var = "black")

```

Analizzando le coordinate per i singoli attributi si riesce a capire i 4 cluster rappresentano i 4 tipi di variabili in gioco (Red, Green NIR1, NIR2)

```{r}
# Extract the principal components
dataset.input.pca = data.frame(get_pca_ind(pca.results)$coord)
dataset.pca = cbind(class=dataset$class, dataset.input.pca)

# Draw some feature plot to highlight the level of difficulty to recognise the various classes
featurePlot(x=dataset.input.pca, y=dataset.output, plot="density", scales=list(x=list(relation="free"), y=list(relation="free")), auto.key=list(columns=3), main="Density Feature Plot")
featurePlot(x=dataset.input.pca, y=dataset.output, plot="pairs", auto.key=list(columns=3), main="Pairs Feature Plot")
```


```{r}

# Divide the dataset into trainset and testset
indexes = createDataPartition(dataset.pca$class, p = .7, list = F, times = 1)
temp.trainset = dataset.pca[indexes,]
testset = dataset.pca[- indexes,]
testset.x = testset[, !names(testset) %in% c("class")]
testset.y = testset[, c("class")]

# Divide the trainset to obtain a validationset
indexes = createDataPartition(temp.trainset$class, p = .9, list = F, times = 1)
trainset = temp.trainset[indexes,]
validationset = temp.trainset[-indexes,]
validationset.x = validationset[, !names(validationset) %in% c("class")]
validationset.y =  validationset[, c("class")]
```

```{r}
# Train a SVM model and tune the hyperparameters (cost and gamma)
svm.start_time = Sys.time()

svm.tuning.control = tune.control(sampling = "fix") # To use validationset instead
svm.tuning = tune.svm(class ~ Dim.1 + Dim.2 + Dim.3 + Dim.4, data = trainset, kernel = "radial", gamma = 10^(-3:1), cost = 10^(-2:1), validation.x = validationset.x, validation.y = validationset.y, tunecontrol = svm.tuning.control, probability = TRUE)

svm.end_time = Sys.time()
sprintf("Time required for svm training:")
svm.end_time - svm.start_time

```

```{r}
plot(svm.tuning)
```


```{r}
# Determine some metrics for each class
svm.tuned = svm.tuning$best.model
svm.tuned.prediction = predict(svm.tuned, testset.x)
svm.tuned.result = confusionMatrix(svm.tuned.prediction, testset.y, mode = "prec_recall")

sprintf("Tuned SVM micro-average global accuracy is %2.1f%%", svm.tuned.result$overall["Accuracy"] * 100)
print(svm.tuned.result$table)
```

```{r}
# Train a NNET model and tune the hyperparameters (size and decay)
nnet.start_time = Sys.time()

nnet.tuning.control = tune.control(sampling = 'fix')
nnet.tuning = tune.nnet(class ~ Dim.1 + Dim.2 + Dim.3 + Dim.4, data = trainset, size = c(2:4), validation.x = validationset.x, validation.y = validationset.y, decay = c(0, 2^-2:1), probability = TRUE)

nnet.end_time = Sys.time()
sprintf("Time required for svm training:")
nnet.end_time - nnet.start_time
```
```{r}
plot(nnet.tuning)
```


```{r}
# Determine some metrics for each class
nnet.tuned = nnet.tuning$best.model
nnet.tuned.prediction = factor(predict(nnet.tuned, testset.x, type = "class"), levels = class.names)
nnet.tuned.result = confusionMatrix(nnet.tuned.prediction, testset.y, mode = "prec_recall")

sprintf("Tuned NNET micro-average global accuracy is %2.1f%%", nnet.tuned.result$overall["Accuracy"] * 100)
print(nnet.tuned.result$table)
```


## Confronto modelli


### Calcolo metriche
In questa sezione vengono calcolate e confrontate metriche di base come sensitivity, specificity e F-scores.

```{r}
metrics_of_interest = c("Precision", "Specificity", "Recall", "F1")
```


```{r}
svm.macro.fscore = Reduce('+', svm.tuned.result$byClass[,"F1"]) / length(class.names)
svm.micro.fscore = Reduce('+', svm.tuned.result$byClass[,"F1"] * svm.tuned.result$byClass[,"Prevalence"])

sprintf("SVM macro F-Score: %1.3f", svm.macro.fscore)
sprintf("SVM micro F-Score: %1.3f", svm.micro.fscore)

print(svm.tuned.result$byClass[, metrics_of_interest])
```

```{r}
# Determine some metrics for each class
nnet.macro.fscore = Reduce('+', nnet.tuned.result$byClass[,"F1"]) / length(class.names)
nnet.micro.fscore = Reduce('+', nnet.tuned.result$byClass[,"F1"] * nnet.tuned.result$byClass[,"Prevalence"])
sprintf("NNET macro F-Score: %1.3f", nnet.macro.fscore)
sprintf("NNET micro F-Score: %1.3f", nnet.micro.fscore)
sprintf("Tuned NNET metrics by class:")
nnet.tuned.result$byClass[, metrics_of_interest]
```

```{r}
metrics.comparison = svm.tuned.result$byClass[, metrics_of_interest] - nnet.tuned.result$byClass[, metrics_of_interest]
print("Metrics compared, positive values favor SVM")
print(metrics.comparison)

```

### Curve ROC
```{r}
# Predicts the most probable class class for each instance
svm.tuned.predictions = predict(svm.tuned, testset.x, probability = TRUE)

# Extracts, for each class, the probability of every instance of being of that class 
svm.tuned.predictions.probs = attr(svm.tuned.predictions, "probabilities")

# Extracts, for each class, the probability of every instance of being of that class
nnet.tuned.predictions = predict(nnet.tuned, testset.x, probability = TRUE)

# Creates the dataframe to be used for multiroc
testset.multiroc = data.frame(matrix(ncol=18, nrow=0))

for (i in 1:length(testset.y)) {
    row = matrix(0, 1, 6)
    # sets 1 to the correct class 
    row[testset.y[i]] = 1
    for (j in 1:6)
        # concatenates the probabilities of the svm
        row = c(row, svm.tuned.predictions.probs[i, class.names[j]])
    # concatenates the probabilities of the neural net
    row = c(row, nnet.tuned.predictions[i,])
    testset.multiroc = rbind(testset.multiroc, row)
}

# Formats the dataframe to be suitable for the multi_roc function
colnames(testset.multiroc) = c('red_soil_true', 'cotton_crop_true', 'grey_soil_true', 'damp_grey_soil_true', 'vegetation_true', 'very_damp grey_soil_true', 'red_soil_pred_SVM', 'cotton_crop_pred_SVM', 'grey_soil_pred_SVM', 'damp_grey_soil_pred_SVM', 'vegetation_pred_SVM', 'very_damp grey_soil_pred_SVM', 'red_soil_pred_NN', 'cotton_crop_pred_NN', 'grey_soil_pred_NN', 'damp_grey_soil_pred_NN', 'vegetation_pred_NN', 'very_damp grey_soil_pred_NN')

svm.nnet.multiroc = multi_roc(testset.multiroc, force_diag=T)
print("Per class AUC values:")
print(unlist(svm.nnet.multiroc$AUC))
```


```{r}
# MultiROC plotting
n_method <- length(unique(svm.nnet.multiroc$Methods))
n_group <- length(unique(svm.nnet.multiroc$Groups))

# changes the format of results to a ggplot2 friendly format
for (i in 1:n_group) {
      res_df <- data.frame(Specificity= numeric(0), Sensitivity= numeric(0), Group = character(0), AUC = numeric(0), Method = character(0))
      for (j in 1:n_method) {
        temp_data_1 <- data.frame(Specificity=svm.nnet.multiroc$Specificity[[j]][i],
                                  Sensitivity=svm.nnet.multiroc$Sensitivity[[j]][i],
                                  Group=tools::toTitleCase(gsub("_", " ", unique(svm.nnet.multiroc$Groups)[i])),
                                  AUC=svm.nnet.multiroc$AUC[[j]][i],
                                  Method = unique(svm.nnet.multiroc$Methods)[j])
        colnames(temp_data_1) <- c("Specificity", "Sensitivity", "Group", "AUC", "Method")
        res_df <- rbind(res_df, temp_data_1)

      }
      
      print(ggplot2::ggplot(res_df, ggplot2::aes(x = 1-Specificity, y=Sensitivity)) + ggplot2::geom_path(ggplot2::aes(color = Group, linetype=Method)) + ggplot2::geom_segment(ggplot2::aes(x = 0, y = 0, xend = 1, yend = 1), colour='grey', linetype = 'dotdash') + ggplot2::theme_bw() + ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5), legend.justification=c(1, 0), legend.position=c(.95, .05), legend.title=ggplot2::element_blank(), legend.background = ggplot2::element_rect(fill=NULL, size=0.5, linetype="solid", colour ="black")))
    
}

```

## Clustering

```{r}
source("clustering_metrics.R")
dataset.clustering = dataset.input.pca
labels = dataset.pca[1]

# dataframe to vector conversion
labels.list <- c()
labels.list = match(as.factor(labels$class), class.names)
```

### Tendency Measures

```{r}
# Calculate Hopkins statistic for CLustering tendency
set.seed(1)
hopkins(dataset.clustering, n = nrow(dataset.clustering) - 1)

```

```{r}
# Since the entire dataset is too big to be allocated for VAT result, only a portion of it has been given in input
indexes = createDataPartition(dataset.pca$class, p = .2, list = F, times = 1)
partition = dataset.pca[indexes,]
partition.vat = partition[, !names(testset) %in% c("class")]

# VAT - Visual Assessment of cluster Tendency
fviz_dist(dist(partition.vat), show_labels = FALSE)

```


### Quality Measures

# Scelta dei parametri di seed e nstart
```{r}
# Il valore ottimale per nstart e' stato individuato in 5, che garantisce un'ottima misura di stabilita' dell'algorimto, con una bassa influenza del seed iniziale rispetto alla suddivisione in clusterr del dataset

# Di seguito e' mostrato come i diversi valori di seed (scelti in maniera casuale), non influenzino la misura di silhouette
Z <- array(sample(1:100, 20, replace=T), c(20, 1))
silhouettes <- c()
for(i in 1:20) {
    set.seed(Z[i])
    fit = kmeans(dataset.clustering, 3, nstart=5)
    ss = silhouette(fit$cluster, dist(dataset.clustering))
    silhouettes = append(silhouettes, mean(ss[, 3]))
    # print(sprintf("Seed %2d: %f", Z[i], mean(ss[, 3])))
}
plot(Z, silhouettes, type="b", main = "Silhouette plot with different random seeds", xlab="seeds")
set.seed(1)
```

```{r}
set.seed(1)
nk = 2:12

# calculates average silhouette values for an ascending value of k
SW = sapply(nk, function(k){cluster.stats(dist(dataset.clustering), 
                              kmeans(dataset.clustering, centers=k, nstart=5)$cluster)$avg.silwidth})
#SW
plot(nk, SW, type="l", xlab="number of clusters", ylab="average silhouette")
```

### Evaluation al variare di k
Evidenzia la misura di similarita' tra i cluster e le classi del dataset, al variaaare del parametro k
Il valore piu' alto e' ottenuto con k=6, in linea con le rilevazioni precedenti
```{r}
similarity = c()

# Calculates the similarity between clusters and targets for an ascending value of k
for (k in 2:6) {
    fit = kmeans(dataset.clustering, k, nstart = 5)
    eval = cluster.evaluation(labels.list, fit$cluster)
    similarity = append(similarity, eval)
}
plot(2:6, similarity, type="b", main = "Similarity plot for ascending K", xlab="K")
```


### K=6
Pur ottenendo una misura minore di Silhouette in confronto a k=3, e' stato comunque ritenuto opportuno analizzare i risultati ottenuti tramite l'esecuzioe con parametro k=6, che rappresenta il numero delle classi, e di conseguenza porta ad una maggiore misura di similarita' tra i cluster rilevati da k-means e l'effettiva suddivisione in classi del dataset
```{r}
fit = kmeans(dataset.clustering, 6, nstart=5)

# calculates silhouette values for k=6
kms = silhouette(fit$cluster, dist(dataset.clustering))
plot(kms, col=1:6, border=NA, main = "Silhouette Plot for k=6")

```

```{r}
# Dissimilarity Matrix
dissplot(dist(dataset.clustering), labels=fit$cluster, options=list(main="Kmeans clustering with k=6"))
```

```{r}
# Evaluation
sprintf("Similarity with (k=6): %f", cluster.evaluation(labels.list, fit$cluster))
```

```{r}
# visualization
fviz_cluster(fit, dataset.clustering, ellipse.type = "norm", axes=c(1,2), main="Dimensions: 1 & 2", 
             geom="point")
fviz_cluster(fit, dataset.clustering, ellipse.type = "norm", axes=c(1,3), main="Dimensions: 1 & 3", 
             geom="point")
fviz_cluster(fit, dataset.clustering, ellipse.type = "norm", axes=c(1,4), main="Dimensions: 1 & 4", 
             geom="point")

fviz_cluster(fit, dataset.clustering, ellipse.type = "norm", axes=c(2,3), main="Dimensions: 2 & 3", 
             geom="point")
fviz_cluster(fit, dataset.clustering, ellipse.type = "norm", axes=c(2,4), main="Dimensions: 2 & 4", 
             geom="point")

fviz_cluster(fit, dataset.clustering, ellipse.type = "norm", axes=c(3,4), main="Dimensions: 3 & 4", 
             geom="point")
```

```{r}
# Confusion Matrix

confusion_matrix = clustering_confusion_matrix(fit$cluster, labels.list, 6, 6)
colnames(confusion_matrix) = class.names
rownames(confusion_matrix) = 1:6
confusion_matrix
```

```{r}
for (i in 1:6) {
    pie_title = sprintf("Cluster %d Distribution", i)
    pie(confusion_matrix[i,], main=pie_title, labels=class.names, col=rainbow(6))
}
```

```{r}
#
for (i in 1:6) {
    pie_title = sprintf("Class %s Distribution", class.names[i])
    pie(confusion_matrix[,i], main=pie_title, labels=1:6, col=rainbow(6))
}
```
```{r}
source("clustering_metrics.R")

precisions = c()
recalls = c()
fmeasures = c()

for (i in 1:6) {
    precision = clustering_precision(confusion_matrix, i)
    recall = clustering_recall(confusion_matrix, i)
    fmeasure = (2*precision*recall) / (precision+recall)
    
    precisions = rbind(precisions, precision)
    recalls = rbind(recalls, recall)
    fmeasures = rbind(fmeasures, fmeasure)
}

km_metrics = cbind(precisions, recalls, fmeasures)

rownames(km_metrics) = class.names
colnames(km_metrics) = c('precision','recall','fmeasure')

accuracy = clustering_accuracy(confusion_matrix)
sprintf("K-means (K=6) micro-average global accuracy is %2.1f%%", accuracy * 100)

sprintf("Per class metrics:")
km_metrics
```
